Nested sampling with two or more objective functions

### 2015-10-09 Paper/ms.tex
Call the scalars L

In some inference and
statistical mechanics problems, there are two or more scalar functions of
$\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
we obtain testable information that fixes the expected values of two scalar
functions of $\xx$, $S_1(\xx)$ and $S_2(\xx)$. It is well known that the
updated probability distribution that takes into account the constraints is
of the ``canonical'' form:

\begin{eqnarray}
p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1S_1(\xx)+\beta_2S_2(\xx)\right]}
{Z(\beta_1, \beta_2)}
\end{eqnarray}

If we were only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
constant and by running standard Nested Sampling with ``likelihood''
$L(\xx) = \exp\left[0.3S_1(\xx) + 0.7S_2(\xx)\right]$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the normalisation $Z(\beta_1, \beta_2)$, called the
partition function. This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, we require the method
to satisfy the following requirements:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$S_1(\xx)$ and $S_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should try to move through a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$S_1$ and $S_2$, i.e. it should only depend on rankings of $S_1$ and $S_2$
values and not the values themselves.
\end{enumerate}

\section{Draft algorithm}

The algorithm works as follows.
Start with $N$ particles drawn from $\pi$.
Define the {\it corner count} of particle $i$ by
\begin{eqnarray}
C_i &=& \sum_{j=1}^N \mathds{1}\left(S_1^j < S_1^i, S_2^j < S_2^i\right)
\end{eqnarray}

>>>>

In some inference and
statistical mechanics problems, there are two or more scalar functions of
$\xx$ that are relevant. Suppose our prior is $\pi(\xx)$ as before, and
we obtain testable information that fixes the expected values of two scalar
functions of $\xx$, $L_1(\xx)$ and $L_2(\xx)$. It is well known that the
updated probability distribution that takes into account the constraints is
of the ``canonical'' form:

\begin{eqnarray}
p(\xx; \beta_1, \beta_2) &=& \frac{\pi(\xx)\exp\left[\beta_1L_1(\xx)+\beta_2L_2(\xx)\right]}
{Z(\beta_1, \beta_2)}
\end{eqnarray}

If we were only interested in a single canonical distribution, for example
with $\beta_1 = 0.3$ and $\beta_2 = 0.7$, we could estimate its normalising
constant and by running standard Nested Sampling with ``likelihood''
$L(\xx) = \exp\left[0.3L_1(\xx) + 0.7L_2(\xx)\right]$. However, usually we
are interested in a range of values for $\beta_1$ and $\beta_2$, and we
want to know the normalisation $Z(\beta_1, \beta_2)$, called the
partition function. This work describes
progress towards solving this class of problems while maintaining the benefits
of Nested Sampling, such as the ability to cope with first-order phase
transitions.

@@

We seek an algorithm for computing the partition function
$Z(\beta_1, \beta_2)$ for a range of values of the $\beta$s, from a single
run. To be considered a variant of Nested Sampling, we require the method
to satisfy the following requirements:
\begin{enumerate}
\item The algorithm should begin with $N$ points drawn from the prior $\pi(\xx)$.
\item The algorithm should seek to explore regions where the values of
$L_1(\xx)$ and $L_2(\xx)$ are greater than what would be expected from the
prior.
\item The algorithm should try to move through a sequence of probability
distributions proportional to $\pi$, but restricted to smaller and smaller
domains for which we can approximately measure the enclosed prior mass.
\item The algorithm should be invariant to monotonic transformations of
$L_1$ and $L_2$, i.e. it should only depend on rankings of $L_1$ and $L_2$
values and not the values themselves.
\end{enumerate}

\section{Proposed algorithm}

The algorithm works as follows.

\begin{enumerate}
\item Generate $N$ particles $\{\xx_1, ..., \xx_N\}$ from $\pi(\xx)$.\\
\item Calculate the {\it corner count} (Equation~\ref{eqn:corner_count})
of each particle.\\
\end{enumerate}

The corner count of particle $i$ is defined by
\begin{eqnarray}
C_i &=& \sum_{j=1}^N \mathds{1}\left(L_1(\xx_j) < L_1(\xx_i),
L_2(\xx_j) < L_2(\xx_i)
\right).\label{eqn:corner_count}
\end{eqnarray}
